The cost of undisturbed landscapes

### 2020-04-28 projects/asparagus/doc/slides.tex
updates asparagus slides

        \begin{textblock*}{5.75cm}
            (1.0cm,2cm)
            Consider a renewable energy technology $\omega$ that
            \begin{itemize}
                \item is \emph{system cost} minimizing, i.e.\ it saves cost relative to its best alternative $\mu$, but
                \item comes at \emph{social cost}, e.g. due to unaccounted negative external effects
            \end{itemize}
            \medskip

            A social planner should substitute $\omega$ with $\mu$ up to the point where the increase in system cost
            equals the reduction in harm (from $\omega$'s negative externality unaccounted in system cost)
        \end{textblock*}

        \begin{textblock*}{7.5cm}
            (7.5cm,2cm)
            \begin{tikzpicture}
                \begin{axis}[
                    width=\textwidth,
                    clip=false,
                    axis lines=left,
                    ticks=none,
                    xlabel={$\omega$ substituted with equivalent $\mu$},
                    ylabel={cost},
                    domain=0.5:1.75,
                    xmin=0.5, xmax=2,
                    ymin=0, ymax=4,
                    samples=200
                ]
                    \addplot[blue, thick] (x, 2*x^1.1-0.933);
                    \node at (axis cs: 1.925,3) [blue, align=left] {additional\\ system\\ cost};

                    \addplot[bokugreen, thick] (x, -2*x^0.8+4);
                    \node at (axis cs: 2.0,1) [bokugreen, align=left]{harm \\(from RET\\ externality)};
                \end{axis}
            \end{tikzpicture}
        \end{textblock*}
    \end{frame}

@@

    \begin{frame}
        \frametitle{The case of Austria}
        \begin{itemize}
            \item Policy goal: \SI{100}{\percent}\footnote[frame]{excluding system services and industry own consumption.
            At current levels this equals 10\% of consumption, i.e. actual target is around 90\%.} of electricity demand from
            domestic renewable sources on annual balance by 2030
            \begin{itemize}
                \item[$\rightarrow$] Ideal frame for studying substitution / opportunity cost
            \end{itemize}
            \item necessarily turns Austria into a net exporter of electricity
            \item Official expectation: additional generation of \SI{27}{\tera\watt\hour} sufficient
        \end{itemize}
        \begin{table}
            \centering
            \begin{tabular}{| l | r | r | r | r | r |}
                \hline
                & 2018 & 2018\footnote[frame]{meteorological conditions as in 2016} & Policy & 2030 & 2030 \\
                & [GW] & [TWh] & [TWh] & [TWh] & [GW] \\ \hline \hline
                Solar PV & 1.44 & 1.23 & +11 & 12.23 & 14.27 \\ \hline
                Wind (onshore) & 3.05 & 6.14 & +10 & 16.14 & 8 \\ \hline
                Hydro (run-of-river) & 5.72 & 28.34 & +5 & 33.34 & 6.73 \\ \hline
                Biomass & $\sim 1$ & 4.78 & +1 & 5.78 & $<2$ \\ \hline
            \end{tabular}
        \end{table}
    \end{frame}

@@

\medskip
\begin{center}
    \emph{medea} is available on \textcolor{blue}{\href{https://github.com/inwe-boku/medea}{github.com/inwe-boku/medea}}
    under an open MIT license
\end{center}
\end{frame}
\note{
medea is a partial-equilibrium model of the coupled electricity and district heating sectors.
Operation of and investment in conventional and renewable power plants, storages and transmission lines are determined simultaneously for one year.
Results include hourly dispatch, investment, inter-zonal trade, electricity and heat prices, profits, and emissions of air pollutants.
The model accounts for various technical characteristics of power systems, such as co-generation of heat and power, balancing requirements, 
}

\section{Results}

>>>>

        \begin{textblock*}{5.75cm}
            (1.0cm,2cm)
            Consider a renewable energy technology $T$ that
            \begin{itemize}
                \item is deployed at the \emph{system cost} minimizing level $T^\ast$, i.e.\ it saves system cost
                relative to its best alternative $P$, but
                \item comes at \emph{social cost}, e.g. due to unaccounted negative external effects
            \end{itemize}
            \medskip

            A social planner should substitute $T$ with $P$ up to the point where the increase in system cost
            equals the reduction in harm (from $T$'s negative externality which is unaccounted for in system cost)
        \end{textblock*}

        \begin{textblock*}{7.5cm}
            (7.5cm,2cm)
            \begin{tikzpicture}
                \begin{axis}[
                    width=\textwidth,
                    clip=false,
                    axis lines=left,
                    ticks=none,
                    xlabel={$\Phi= \frac{T^\ast-T}{T^\ast}$},
                    ylabel={cost},
                    domain=0.5:1.75,
                    xmin=0.5, xmax=2,
                    ymin=0, ymax=4,
                    samples=200
                ]
                    \addplot[blue, thick] (x, 2*x^1.1-0.933);
                    \node at (axis cs: 2.05,2.95) [blue, align=left] {additional\\ system cost};

                    \addplot[bokugreen, thick] (x, -2*x^0.8+4);
                    \node at (axis cs: 1.9,0.925) [bokugreen, align=left]{harm};
                    \node at (axis cs: 0.5,-0.25) []{$0$};
                    \node at (axis cs: 1.925,-0.25) []{$1$};
                \end{axis}
            \end{tikzpicture}
        \end{textblock*}
    \end{frame}

@@

    \begin{frame}
        \frametitle{The case of Austria}
        \begin{itemize}
            \item Policy goal: \SI{100}{\percent}\footnote[frame]{excluding system services and industry own consumption.
            At current levels this equals 10\% of consumption, i.e. actual target is around 90\%.} of electricity demand from
            domestic renewable sources on annual balance by 2030
            \begin{itemize}
                \item[$\rightarrow$] Ideal frame for studying substitution / opportunity cost
            \end{itemize}
            \item necessarily turns Austria into a net exporter of electricity
            \item Official expectation: additional generation of \SI{27}{\tera\watt\hour} sufficient
        \end{itemize}
        \begin{table}
            \centering
            \begin{tabular}{lrrrrr}
                %\hline
                & 2018 & 2018\footnote[frame]{meteorological conditions as in 2016} & Policy & 2030 & 2030 \\
                & [GW] & [TWh] & [TWh] & [TWh] & [GW] \\ \hline \hline
                Solar PV & 1.44 & 1.23 & +11 & 12.23 & 14.27 \\ \hline
                Wind (onshore) & 3.05 & 6.14 & +10 & 16.14 & 8 \\ \hline
                Hydro (run-of-river) & 5.72 & 28.34 & +5 & 33.34 & 6.73 \\ \hline
                Biomass & $\sim 1$ & 4.78 & +1 & 5.78 & $<2$ %\\ \hline
            \end{tabular}
        \end{table}
    \end{frame}

@@

\medskip
\begin{center}
    \emph{medea} is available on \textcolor{blue}{\href{https://github.com/inwe-boku/medea}{github.com/inwe-boku/medea}}
    (MIT license)
\end{center}
\end{frame}
    \note{
        medea is a partial-equilibrium model of the coupled electricity and district heating sectors.
        Operation of and investment in conventional and renewable power plants, storages and transmission lines are determined simultaneously for one year.
        Results include hourly dispatch, investment, inter-zonal trade, electricity and heat prices, profits, and emissions of air pollutants.
        The model accounts for various technical characteristics of power systems, such as co-generation of heat and power, balancing requirements,
    }

    \section{Results}
    \begin{frame}
        \frametitle{The cost of undisturbed landscapes}

        \begin{textblock*}{10cm}
            (0.75cm, 2.25cm) % {block width} (coords)
            \includegraphics[width=\textwidth, keepaspectratio]{./figures/undisturbed_base_share.pdf}
        \end{textblock*}

                \vspace{3.75cm}
                \quad wind power and solar PV \\ \quad are substitutes
            \end{small}
        \end{textblock*}
    \end{frame}
